Synthesizing Symmetric Lenses

### 2018-07-03 papers/SymmetricLenses/main.tex
More comments

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

To answer our first research question, we analyze the runtime of our algorithm,
and compare it to the existing algorithm for synthesizing bijective lenses.

@@

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{ Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
    symmetric synthesis algorithm without using a library of existing lenses.
    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete most of the benchmarks without using compositional synthesis. }
  \label{fig:times}
\end{figure}

@@

\subsection{Heuristic Usefulness}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in 3 different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \NoTPOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
                but users cannot tune when \RXSearch returns using a termination parameter.\\
  \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
  \NMOpt{}: & \GreedySynth does not have any notion of cost, and searches for
              any lens that satisfies the specification. \\
\end{tabulary}

Figure~\ref{fig:heuristic} summarizes the result of this experiment.  We find
that \afm{TODO}.
% end evaluation

@@

These components only make sense for synchronization when they satisfy
round-tripping laws.  This guarantees that if an identity edit is propagated, no
data gets changed.
\begin{equation}
  \tag{\PutRL}
  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'')} = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
\end{equation}

@@

Similar to the previous case.
\end{enumerate}
\end{proof}
\end{document}

>>>>

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

To answer our first research question, we measure the runtime of our algorithm
and compare it to the existing algorithm for synthesizing bijective
lenses.\bcp{Note that this comparison doesn't, directly answer the question
  ``Is it fast enough.''}

@@

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{ Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
    symmetric synthesis algorithm without using a library of existing lenses.
    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete most of the benchmarks without using compositional
    synthesis. \bcp{The interesting part of the graph is all squished
      against the Y axis.  Can we make the X axis logarithmic?}}
  \label{fig:times}
\end{figure}

@@

\subsection{The Effect of Our Heuristic}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in 3 different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \NoTPOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
                but users cannot tune when \RXSearch returns using a termination parameter.\\
  \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
  \NMCCOpt{}: & The lens cluster that contains is identified through \SSOpt{}. \GreedySynth is then called on that pair of SREs. \\
  \NMOpt{}: & Neither \GreedySynth nor \RXSearch have any notion of cost; they
              both merely search for
              any lens that satisfies the specification. 
\end{tabulary}

\begin{figure}
  \includegraphics{generated-graphs/metrics_importance}
  \caption{Number of benchmarks that synthesize the correct lens by a given
    algorithm. \NoTPOpt{} is the full symmetric synthesis algorithm, but
    termination parameters are ignored. \FCOpt{} provides a notion of cost to
    \GreedySynth, but once a satisfying lens is greedily found, that lens is
    returned. \NMCCOpt{} identifies the lens cluster in which the desired 
    lens lies.  \GreedySynth{} is then run on that cluster with no notion of
    cost. \NMOpt{} removes all notion of cost, and merely finds the first
    satisfying lens.\bcp{Am I reading this wrong, or does even the best
      algorithm fail to find the correct lens something like 20\% of the
      time?!?!?!}} 
  \label{fig:metric}
\end{figure}

Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
even when provided with the correct lens cluster, without a notion of cost, we
would need constrain our search space more with additional examples to find the
correct lens.

Our termination condition is also quite important. When running in \FCOpt{},
only 5 lenses are found. The first cluster in which a lens is found is rarely
the correct cluster. However, our termination condition is not perfect.
Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
Without termination parameters, our algorithm finds the correct lens for 36 of
our 45 benchmarks, 9 of them need help in the form of a termination parameter
for finding the correct lens.
% end evaluation

@@

These components only make sense for synchronization when they satisfy
round-tripping laws.  This guarantees that if an identity edit is propagated, no
data gets changed.
\begin{equation}
  \tag{\PutRL}
  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'''')} = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
\end{equation}

@@

Similar to the previous case.
\end{enumerate}
\end{proof}

\begin{theorem*}
  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
    \Regex}{\String}$ and $\LanguageOf{\Regex} =
  \LanguageOf{\ToDNFRegexOf{\Regex}}$
\end{theorem*}
\begin{proof}
By \sam{COPY PASTE PROOF FROM BIJECTIVE LENS PAPER} $\LanguageOf{\Regex} =
  \LanguageOf{\ToDNFRegexOf{\Regex}}$, so we will prove the result $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
    \Regex}{\String}$.
\begin{enumerate}
\item
$(S = s)$

By definition, $\Downarrow s = \langle ([s],1) \rangle$ and $P_{\langle ([s], 1) \rangle}(s') = 
\begin{cases}
1 & \text{if } s = s'\\
0 & \text{otherwise}
\end{cases}$ hence $P_{\langle ([s],1) \rangle}(s') = P_s(s')$.
\item
$(S = \varnothing)$

By definition, $\Downarrow \varnothing = \langle \rangle$ and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
\item
$(S = S^{*p})$

By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $. By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$, hence
\begin{align*}
P_{[\epsilon \cdot (\Downarrow S)^{*p}] \cdot \epsilon}(s) &= P_{(\Downarrow S)^{*p}}(s)\\
&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{(\Downarrow S)}(s_i)\\
&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{S}(s_i)\\
&= P_{S^{*p}}(s)
\end{align*}
\item
$(S = S_1 \cdot S_2)$

By definition, $\Downarrow (S_1 \cdot S_2) = (\Downarrow S_1) \odot (\Downarrow S_2)$. By the induction hypothesis, $P_{S_1} = P_{(\Downarrow S_1)}$, and $P_{S_2} = P_{(\Downarrow S_2)}$. Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$. Then
\begin{align*}
P_{S_1}(s) &= \sum_{i=1}^n p_i P_{SQ_i}(s)\\
P_{S_1}(s) &= \sum_{j=1}^m q_i P_{TQ_j}(s)\\
\end{align*}
hence
\begin{align*}
P_{S_1 \cdot S_2}(s) &= \sum_{s_1 \cdot s_2 = s} P_{S_1}(s_1)* P_{S_2}(s_2)\\
&= \sum_{s_1 \cdot s_2 = s} \left(\sum_{i=1}^n p_i P_{SQ_i}(s_1)\right)* \left(\sum_{j=1}^m q_i P_{TQ_j}(s_2)\right)\\
&= \sum_{s_1 \cdot s_2 = s} \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j) * (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))
\end{align*}
By definition,
$\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
      \DNFSep
      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
      \ldots$\\
      $\DNFSep
      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
      \ldots \DNFSep
      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$

where
$$\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$$
Hence
$$P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s) = \sum_{i=1}^n \sum_{j=1}^m(p_i * q_j) * P_{SQ_i \odot_{SQ} TQ_j}(s)$$
Since $P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(s) = \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = s} \prod_{i=1}^n P_{A_i}(s'i)$, then
\begin{align*}
P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n] \odot_{SQ} [t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(s) &= \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n \cdot t_0 \cdot t'_1 \cdot \ldots t'_m \cdot t_m =s } \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
&= \sum_{a \cdot b = s} \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
& \sum_{a \cdot b = s} \left(\sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \prod_{i=1}^n P_{A_i}(s'_i)\right) \left(\sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{j=1}^m P_{B_j}(t'_j)\right)\\
&= \sum_{a \cdot b = s}P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(a) * P_{[t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(b)
\end{align*}
In other words, $P_{SQ \odot_{SQ} TQ}(s) = \sum_{a \cdot b = s}P_{SQ}(a) * P_{TQ}(b)$. Hence, 
\begin{align*}
P_{S_1 \cdot S_2}(s) &= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} P_{SQ_i \odot_{SQ} TQ_j}(s)\\
&= P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s)\\
&= P_{(\Downarrow S_1) \odot (\Downarrow S_2)}(s)
\end{align*}
which is what we wanted to show.
\item
$(S = S_1 \; |_p \; S_2)$

By definition $\Downarrow (S_1 \; |_p \; S_2) = (\Downarrow S_1) \oplus_p (\Downarrow S_2)$. By the induction hypothesis $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_2} = P_{\Downarrow S_2}$. Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$. By definition,

  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
  hence
  \begin{align*}
  P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s) &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle \oplus_p \langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s) \\
  &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n) \; | \; (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
  &= \sum_{i=1}^m (p * p_i) P_{SQ_i}(s) + \sum_{j=1}^m ((1-p) * q_j) P_{TQ_j}(s)\\
  &= p * \sum_{i=1}^m p_i P_{SQ_i}(s) + (1-p)* \sum_{j=1}^m q_j * P_{TQ_j}(s)\\
&= p * P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle}(s) + (1-p) * P_{\langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
&= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s) 
  \end{align*}

  Since $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_1} = P_{\Downarrow S_1}$, then 
  \begin{align*}
  P_{S_1}(s) &= \sum_{i=1}^n p_i * P_{SQ_i}(s)\\
  P_{S_2}(s) &= \sum_{j=1}^m q_m * P_{TQ_j}(s)
  \end{align*}
  hence
  \begin{align*}
  P_{S_1 \; |_p \; S_2}(s) &= p * (P_{S_1}(s)) + (1-p) * (P_{S_2}(s))\\
  &= p \left(\sum_{i=1}^n p_i * P_{SQ_i}(s)\right) + (1-p)\left(\sum_{j=1}^m q_m * P_{TQ_j}(s)\right)\\
  &= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s)\\
  &= P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s)
  \end{align*}
  which is what we wanted to show. This completes the proof.
\end{enumerate}
\end{proof}
\begin{theorem*}
  \label{thm:correct_entropy}
  If $\Regex$ is unambiguous and contains no empty subcomponents,
  $\EntropyOf{\Regex}$ is the entropy of $\mathcal{L}(\Regex)$.
\end{theorem*}
\begin{proof}
Given a discrete set $X$ and a probability distribution $P$ on $X$, we define the entropy $H(X)$ of $X$ by
$$H(X) = -\sum_{x \in X}P(x)\log_2{P(x)}$$
We prove the theorem by induction on $S$.
\begin{enumerate}
\item
($S = s$)

Then $\mathbb{H}(s) = 0 = H(\mathcal{L}(s)) = H(\{s\})$
\item
$(S = S^{*p})$

Since $S^{*p}$ is unambiguous, then $S$ is also anambiguous, and for each $s \in \mathcal{L}(S^{*p})$, there exist unique $t_1, \ldots, t_n \in \mathcal{L}(S)$ such that $s = t^s_1 \cdot \ldots \cdot t^s_{n_s}$. By the induction hypothesis, $H(\mathcal{L}(S)) = \mathbb{H}(S)$, thus
\begin{align*}
H(\mathcal{L}(S^{*p})) &= - \sum_{s \in \mathcal{L}(S^{*p})} P_{S^{*p}}(s) \log_2 P_{S^{*p}}(s)\\
&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(p^n * (1 - p) * \prod_{i=1}^n P_S(s_i)\right)\\
&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
&= -(1-p) * \log_2 p \sum_n n * p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)
-(1-p) * \log_2 (1 - p) \sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \\
&-(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\\
&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
-(1-p)\sum_n p^n H(\mathcal{L}(S^n))\\
&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
-(1-p)\sum_n p^n n H(\mathcal{L}(S))\\
&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
-\frac{p}{(1-p)} * H(\mathcal{L}(S))\\
&= \frac{p}{(1-p)}(H(\mathcal{L}(S)) - \log_2 p) -\log_2 (1 - p)\\
&= \frac{p}{(1-p)}(\mathbb{H}(S) - \log_2 p) -\log_2 (1 - p)
\end{align*}
\item
($S = S_1 \cdot S_2$)

Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
\begin{align*}
H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
\end{align*}
\item
($S = S_1 \; |_p \; S_2$)

Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$. Then
\begin{align*}
H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
\end{align*}
\end{enumerate}
\end{proof}
\begin{theorem*}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem*}
\begin{proof}
Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
\end{align*}
Define the simple symmetric lens $\lceil \ell \rceil$ by 
\begin{align*}
\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
\end{align*}
We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
\begin{align*}
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
\end{align*}
\end{proof}

\begin{theorem*}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
      \Lens.\PutROf{s}{t} = t}$
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
      \Lens.\PutLOf{t}{s} = s}$
  \end{enumerate}
\end{theorem*}
\begin{proof}
Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$. We want to show that
$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V_{\ell}(s)) = \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
We proceed by induction over $\ell$. 
\begin{enumerate}
\item
($\ell = \IdentityLensOf{S}$)

Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
